### 2020-04-17 PaperDraft.Rmd

---
title: "Inference of seasonal forcings on reproduction numbers during epidemics"
author: "Fengyang Wang, Lu Lu, Ziyao Tian, Yi Xiao Zhang"
output:
  pdf_document: default
  html_notebook: default
---

## Abstract

Amidst the current COVID-19 epidemic, the prediction of its future spread and the length of the epidemic becomes increasingly significant. We investigate the effects of temperature and humidity on the reproductive rate R of COVID-19, as warmer climates in the summer have historically resulted in diminished spread of influenza. Early epidemiology research in COVID-19 have used simple linear models such as ordinary least squares regression to estimate the relationship between R and climate. We estimate the reproductive rate of the virus using a novel Bayesian inference model with MCMC sampling to daily case data of various climate regions. While results are largely inconclusive, they provide us with a weak posterior update that R decreases with temperature and relative humidity.

## Introduction

Exponential growth of the number of symptomatic cases is usually observed in the initial stages of infectious disease epidemics. Typically, the rate of this growth is decomposed into two quantities:

- The serial interval: time between onset of symptoms of two adjacent individuals in a transmission chain.
- The reproductive number, $R$: the expected number of new infections a single symptomatic individual will cause.

Due to inadequate public data to compute the serial interval, we use the estimated serial interval of $4.7$ days with standard deviation $2.9$ days obtained from Nishiura et al. (2020) [1]. The estimation of the latter quantity is important for policy, because infectious disease outbreaks will switch from an exponential growth regime to an exponential decay regime when the reproductive number falls below 1. The goal of mitigation or suppression policies is therefore to effect change on the reproductive number R. Many infectious diseases are believed to be seasonal, including most respiratory viruses, like the SARS-CoV-2 virus implicated in the 2019–2020 coronavirus pandemic. Wang et al. (2020) [3] have analysed how the weather of locations within China correlates with local point estimates of the reproductive number R. However, because their analysis is frequentist in nature, their estimates of the seasonal forcing do not come with any estimate of uncertainty.

We first produce a baseline model, which predicts the reproductive rate using the EpiEstim package, based on previous work by Thompson et al. (2019) [2], and applied to COVID-19 data as in Abbot et al. (2020) [4]. We then apply ordinary least squares regression against climate variables as well as other underlying regional factors, such as population density, GDP per capita, and latitude, similar to the work produced by Wang et al. (2020)[3]. We compare this against our novel Bayesian framework, and evaluate the significance of the explanatory variables in each model.

Since Italy was one of the first countries affected by the epidemic, it has a longer history of case data before intervention and lockdowns took place. Furthermore, due to the many provinces of Italy having diverse climates, we have access to a larger range of temperature and humidity. Therefore, we sourced the Italian government DPC dataset[5], which contains time series incidence for each province since mid February. Using the temperature and humidity data from weather stations sourced from NOAA, we are able to estimate the climate for over $80$ distinct provinces in the Italian COVID-19 dataset.

## Methodology

Thompson et al. (2019) outline a Bayesian framework for estimation of $R(t)$
using symptomatic case counts, decomposed into local and imported cases. The
EpiEstim R package implements ideas from this paper. A simplified summary of
their approach, ignoring the imported cases:

1. Estimation of (discrete) distribution of serial interval from line-list data
   with intervals reflecting date of infection.
2. Estimation of time-varying $R$ from time-bucketed symptomatic case counts.

We developed a method to estimate the strength of various seasonal forcings on
$R$ through MCMC. This involves both substantial deviations from the second part
of the Thompson et al. model and also less less substantial adjustments for the
nature of our data. Our method does not extend the first part of the Thompson et
al. model.

To sanity-check the results of our method, we compare its conclusions against a
point estimate for the seasonal forcings through linear regression with the
EpiEstim estimates of $R$.

### Incidence likelihood of the Thompson et al. model

In its estimation of time-varying $R$, the Thompson et al. model uses the
following likelihood function for the number of new infections at time $t$:

$$ I(t) \space | \space  I(1), \cdots ,I(t-1),\mathbf{w},R(t) \sim \text{Poisson} \left(R(t) \sum_{s=1}^t I(t-s)w_s \right) $$

where

- $I(t)$ is observed data containing the number of new infections at time $t$
- $R(t)$ is a parameter for the reproduction number at time $t$
- $w_s$ denotes the proportion of serial intervals which take on discrete value
  $s$, i.e. $w_s = \operatorname{P}(\mathrm{SI} = s)$.

Because the Thompson et al. model uses a gamma-distributed prior for $R(t)$ (the
conjugate prior for a Poisson likelihood) and because they chose a convenient
family of serial interval distributions, the Bayesian posterior in the Thompson
et al. model can be described by a closed form analytic expression.

### Extension of Thompson et al. for $R$ subject to forcings

Our Bayesian model extends the method of Thompson et al. to multiple locations,
each with its own incidence time series $I_1(t),\cdots,I_n(t)$ such that

$$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$

where:

- $x_i(t)$ is data of observed seasonal factors at time $t$ and location $i$,
  with each feature standardized to mean $0$ and variance $1$
- $\beta$ is a parameter describing the locally linear effect of each seasonal
  forcing
- $\theta(t)$ is a parameter for logarithm of country-wide $R$ independent of
  modeled seasonal forcings
- $\lambda_i$ is a parameter for local variation of $\log R$ independent of
  modeled seasonal forcings
- $\varepsilon_i(t)$ is a parameter for remaining fluctuation in $\log R$ not
  modeled above

Unlike the Thompson et al., our analysis fundamentally estimates the logarithm
of $R$ instead of $R$ itself. This has some practical advantages: a linear model
for $R$ could predict negative values for sufficiently extreme climate
conditions, but negative values for $R$ do not make sense epidemiologically.

We maintain a very similar incidence likelihood function as the Thompson et al.
model, but impose an additional correction for importations:

$$
I_i(t) \space |
\space P_i,x_i(t),I_i(1), \cdots ,I_i(t-1),\mathbf{w},\theta(t),\lambda_i,\mathbf{\beta},\mathbf{\varepsilon}(t),\gamma
\sim \text{Poisson} \left(\gamma P_i + R_i(t) \sum_{s=1}^t I_i(t-s)w_s \right)
$$

where in addition to the parameters described (with $R_i(t)$ given by the
equation above), the following additional data and parameters are introduced:

- $P_i$ is data giving the population of location $i$
- $\gamma$ is a parameter estimating the nationwide rate of imports per capita
per day, on which we impose prior $\gamma \sim \operatorname{Gamma}(2, 10^{-5})$

The only substantative change to this likelihood function, beyond the
decomposition of $R_i(t)$ into the various parameters influencing it, is the
addition of importation risk. The reason for this adjustment is that the dataset
we are using does not distinguish between importations and local cases. Thus,
the Thompson likelihood would be $0$ upon detection of the first case, had we
not modeled the importations separately.

The most important feature of our model is the parameter $\mathbf{\beta}$ which
are the coefficients of the linear model of $\log R$. This feature is assumed to
be global across all locations and times, and thus models inherent properties of
the contagiousness of the outbreak, as opposed to local deviations in space or
time.

To avoid the model ascribing deviations in $R$ to the seasonal forcings without
widespread evidence across locations and times, three other features of our
model are the parameters $\theta(t)$, $\lambda_i$, and $\varepsilon_i(t)$. The
first parameter, $\theta(t)$, describes the logarithm country-wide $R$ before
adjusting for location and climate and local idiosyncratic factors. That is, it
can be thought of as a baseline $R$ value affected simply by the national
circumstances at the given time; such as national policies or human behavioral
changes. This will encapsulate various unmodeled behaviours such as weekends or
interventions. At time zero, our prior belief is that this $R$ will match the
consensus of international estimates for $R_0$, which vary widely but tend to
cluster between $2$ or $3$ [TK: Citation needed?]. Because of the great
uncertainty, there is significant variance in this assumption, so the prior is
quite weak (note that the $\mu \pm 2\sigma$ values, when transformed
exponentially, are below $1$ and above $6$ respectively).

$$
\theta(0) \sim \operatorname{Normal}(\log 2.5, 0.5^2)
$$

The remaining days of $\theta$ are modeled using a simple autoregressive model:

$$
\theta(t+1) \sim \operatorname{Normal}(\theta(t), \theta^\sigma)
$$

where $\theta^\sigma$ is a parameter describing the spread of daily drifts in
$\theta$. The prior on $\theta^\sigma$ is weakly informative:

$$
\theta^\sigma \sim \operatorname{Gamma}(2, 0.01)
$$

The location-specific deviation from the country average in $\log R$ independent
of modeled seasonal forcings is encoded in $\lambda_i$. The existence of this
parameter helps the model explain the large variations of apparent $R$ across
different locations, some of which appear to be affected much more so than
others. The remaining parameter $\varepsilon_i(t)$ encapsulates all other
unexplained deviations across both time and location. All these parameters (and
$\beta$) are given the following weakly informative prior:

$$
\lambda_i, \varepsilon_i(t), \beta_k \sim \operatorname{Normal}(0, 0.5)
$$

Our technique is general and can be used for any epidemic and any combination of
potential seasonal forcings.

As the model is much more complex than the one in the Thompson et al. paper, we
are unable to derive an analytic expression to compute the posterior. Instead,
we use `rstan` to run $4$ chains of MCMC for $2000$ iterations each.

### Using OLS and EpiEstim to check results

The baseline model we build estimates the reproductive rate using a parametric SI model, where we specify the mean and standard deviation of the serial interval. Because we do not have access to line-list SI data for COVID-19, we are unable to effectively estimate the serial interval from data, thus we take the findings of Nishiura et al. (2020), which estimated the mean serial interval to be 4.7 and the standard deviation to be 2.4. We use the model detailed in Cori et al. (2013) [7] for estimating R from bayesian inference on serial interval and time series data, which can be easily implemented using the EpiEstim package.

### Dataset

[TK: retained from proposal, need to modify]

- The Italian government dataset at https://github.com/pcm-dpc/COVID-19 contains the disease incidence for each province. This data contains date of confirmation instead of date of onset of symptoms, and so some procedure may be necessary to correct for this. Abbott et al. (2020) outlines a technique to correct for reporting delays using an exponential distribution of reporting delay fit to line list data; we may use the same distribution.
- The `worldmet` R package enables access to temperature and humidity data, sourced from NOAA.

## Results

## Discussion

### Directions for Future Research

[TK: retained from proposal, need to modify]

- Looking at the climate and infection data in American and Canadian cities as the situation develops further. In particular, compare this with the results in Italy, and observe if the linear assumption between seasonal factors and $R$ holds true across different climate regions. If not, try fitting while assuming a nonlinear relationship between $R$ and seasonal factors.
- Look into statistical simulation models which consider intercity travel and geographic proximity.

## Appendix

### References

[1] Nishiura H, Linton NM, Akhmetzhanov AR. Serial interval of novel coronavirus (2019-nCoV) infections. medRxiv Published Online First: 2020. doi:10.1101/2020.02.03.20019497

[2] R.N. Thompson, J.E. Stockwin, R.D. van Gaalen, J.A. Polonsky, Z.N. Kamvar, P.A. Demarsh, E. Dahlqwist, S. Li, E. Miguel, T. Jombart, J. Lessler, S. Cauchemez, A. Cori. Improved inference of time-varying reproduction numbers during infectious disease outbreaks, Epidemics, Volume 29, 2019, 100356, ISSN 1755-4365, https://doi.org/10.1016/j.epidem.2019.100356.

[3] Wang, Jingyuan and Tang, Ke and Feng, Kai and Lv, Weifeng, High Temperature and High Humidity Reduce the Transmission of COVID-19 (March 9, 2020). http://dx.doi.org/10.2139/ssrn.3551767

[4] (PREPRINT) Abbott, S., Hellewell, J., Munday, J. D., Chun, J. Y., Thompson, R. N., Bosse, N. I., Chan, Y. D., Russell, T. W., Jarvis, C. I., ... & Funk, S. (2020). Temporal variation in transmission during the COVID-19 outbreak. CMMID. Available at CMMID: https://cmmid.github.io/topics/covid19/current-patterns-transmission/global-time-varying-transmission.html

[5] Presidenza del Consiglio dei Ministri - Dipartimento della Protezione Civile, Dati COVID-19 Italia, (2020), GitHub repository, https://github.com/pcm-dpc/COVID-19

[6] “Report 13 - Estimating the Number of Infections and the Impact of Non-Pharmaceutical Interventions on COVID-19 in 11 European Countries.” Imperial College London, www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-13-europe-npi-impact/.

[7] Anne Cori, Neil M. Ferguson, Christophe Fraser, Simon Cauchemez, A New Framework and Software to Estimate Time-Varying Reproduction Numbers During Epidemics, American Journal of Epidemiology, Volume 178, Issue 9, 1 November 2013, Pages 1505–1512, https://doi.org/10.1093/aje/kwt133